---
title: "intermediate script"
author: "Molly Williams"
date: "10/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# .Rprofile

```{r}

library(tidyverse)

install.packages("usethis")

usethis::edit_r_profile(scope = "project") # opens up a new Rprofile script


cowsay::say(what = "hello r4wrds participants!", by = "cow")


```



# .Renviron

Loads variables you want to keep hidden, or allow to be configurable per user

```{r}

# access environmental variables with 
Sys.getenv() # returns all env vars on your machine

# access an existing env var by quoted name
Sys.getenv("USER")

# let's specify our own env vars (custom):
usethis::edit_r_environ(scope = "project")

```


# Create a function

```{r}

inspire_me <- function(animal){

  # get pieces to make link
  repo <- "JakubPetriska/060958fd744ca34f099e947cd080b540"
  csv  <- "raw/963b5a9355f04741239407320ac973a6096cd7b6/quotes.csv"
  
  # silently read dataframe
  suppressMessages(
    quotes  <- readr::read_csv(
      glue::glue("https://gist.githubusercontent.com/{repo}/{csv}")
    )  
  )
  
  # paste together the full quote
  quotes$full_quote  <- paste0(quotes$Quote, " -", quotes$Author)
  
  # make a user-specified animal say the quote
  cowsay::say(sample(quotes$full_quote, 1), by = animal)

}

# have the environmental variable say a quote
inspire_me(Sys.getenv("ANIMAL"))


#can change the animal in .Renviron to any animal 

```


# Modularization of functions


```{r}
# list packages in a vector and load them all
pkgs <- c("readr", "cowsay")
purrr::walk(pkgs, require, character.only = TRUE)

# read quotes from a url
f_read_data <- function(url){
  suppressMessages(
    quotes  <- read_csv(url)  
  )
  return(quotes)
}

# paste the quote to the author
f_preprocess_data <- function(d){
  d$full_quote  <- paste0(d$Quote, " -", d$Author)
  return(d)
}

# print a random animal and a random quote
f_inspire_me <- function(d){
  animals <- names(animals)
  say(sample(d$full_quote, 1), by = sample(animals, 1))
}
```



# Interactive visualizations

```{r}

install.packages("tidygeocoder")

library(tidygeocoder) # geocode our addresses
library(tidyverse)    # wrangle data
library(janitor)      # clean column names
library(glue)         # modern paste() function
library(sf)           # make spatial data
library(mapview)      # interactive maps!
mapviewOptions(fgb = FALSE)


# the url for the Form data 
form_data <- paste0("https://docs.google.com/spreadsheets/d/e/",
                    "2PACX-1vSODxBm_z5Gu8a42C6ZFEa3S5iTbYV-",
                    "qucCGvasGS6c0qFUAml5vSMEgbvI9PYo1HJ20Y_WY62aTAb-",
                    "/pub?gid=1462593645&single=true&output=csv")

# read in url and clean
dat <- read_csv(form_data) %>% # takes the URL above and reads it in
  janitor::clean_names() %>% # comes from janitor package
  dplyr::rename( dining_name = 3, dining_address = 4)

# geocode using Open Street Map (osm) API because it's free
dat_geo <- dat %>%
  geocode(dining_address, method = 'osm', lat = latitude , long = longitude) # looks up the address, returns lat and long, creating new columns for latitude and longitude

# make into sf object so we can map
dat_geo <- dat_geo %>% 
  filter(!is.na(latitude) & !is.na(longitude)) %>% # remove NAs
  st_as_sf(coords = c("longitude", "latitude"), 
           crs = 4326, # spatial projection
           remove = FALSE) # keep lat/lon columns in addition to geometry column being added

# map!
mapview(
  dat_geo, 
  zcol = "comfort_using_r", 
  layer.name = "R comfort level", 
  cex = 6.5
)



```



# Make a fancier map with leaflet

```{r}
install.packages("leafpm")
install.packages("leaflet.extras")

library(leafpm)
library(leaflet)
library(leaflet.extras)
library(htmltools)

# set up our map
m <- leaflet() %>% #call the blank map
  # add tiles or the "basemaps"
  addTiles(group = "OSM") %>% # defaults to Open Street Maps
  addProviderTiles(providers$CartoDB.Positron, group = "Positron") %>% 
  addProviderTiles(providers$Stamen.TonerLite, group = "Toner Lite") %>%
  addCircleMarkers(
    lng = -121.4944, lat = 38.5816, fillColor = "red", color = "black",
    popup = "Sacramento!", group = "Home",
  ) %>% 
  addCircleMarkers(
    data = dat_geo, group = "Food & Drink",
    label = ~htmlEscape(first_name),
    popup = glue(
      "<b>Name:</b> {dat_geo$first_name}<br>
      <b>Food_Name:</b> {dat_geo$dining_name}<br>
      <b>Food_Address:</b> {dat_geo$dining_address}<br>
      <b>R comfort (1-10):</b> {dat_geo$comfort_using_r}"
    )
  )  %>% 
  addLayersControl(
    baseGroups = c("Toner Lite", "Positron", "OSM"),
    overlayGroups = c("Home", "Food & Drink"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>% 
  addMeasure()

m  # Print the map
```


# Network with D3

```{r}

# Libraries
install.packages("igraph")
install.packages("networkD3")

library(igraph)
library(networkD3)

# create a dataset:
data <- tibble(
  from = c(
    "Dam","Dam","Dam", "Dam",
    "River","River","River", "River","River",
    "Canal", "Canal", 
    "Diversion","Diversion", 
    "Reservoir", "Reservoir","Reservoir",
    "Lake","Lake","Lake", "Lake", 
    "Road","Road","Road",
    "Culvert", "Culvert",
    "Fish", "Fish","Fish",
    "Frog","Frog","Frog",
    "MacroInvertebrates","MacroInvertebrates"
  ),
  to = c(
    "River","Reservoir","Canal","Diversion",
    "Lake","Reservoir","Frog","Fish","MacroInvertebrates",
    "Diversion", "Reservoir",
    "Dam", "River",
    "River","Dam","Fish",
    "Fish","Dam","Frog","MacroInvertebrates",
    "Fish","Dam", "Canal",
    "Road", "Dam",
    "Frog", "River","MacroInvertebrates",
    "Fish", "River", "Lake",
    "River", "Lake"
  )
)

# Plot
(p <- simpleNetwork(data, height = "600px", width = "600px", 
                    fontSize = 16, fontFamily = "serif",
                    nodeColour = "darkblue", linkColour = "steelblue",
                    opacity = 0.9, zoom = FALSE, charge = -500))
```




# Plotly: CalEnviroScreen

```{r}
# Load CES data for Sac county
install.packages("cowplot")
library(cowplot)


ces3_sac <- readRDS(here::here("data/ces3_sac.rds"))

mapview(ces3_sac, zcol = "CIscoreP") #map color to the impact of pollution data column



# Static plot of Groundwater threats vs. CES score
(ces_plot <- ces3_sac %>% 
   ggplot(aes(x = gwthreatsP, y = CIscoreP, label = tract)) + 
   geom_point() +
   geom_smooth(method = "gam") +
   cowplot::theme_half_open(font_size = 12) +
   labs(
     title = "CES Score vs. Groundwater Threats in Sacramento County",
     subtitle = "Higher CI Score indicates higher threat/impact",
     x = "Groundwater Threats (percentile)", 
     y = "CI Score (percentile)"
   )
)

# Interactive plotly
install.packages("plotly")
library(plotly)
ggplotly(ces_plot)

```


# Iteration and functional programming

Loops are a building block of functional programming 

## For loops: make iteration very explicit using a counter (i)
```{r}

library(tidyverse)
library(here)

eldorado <- read_csv(here("data", "gwl", "county", "El Dorado.csv"))
placer   <- read_csv(here("data", "gwl", "county", "Placer.csv"))
sac      <- read_csv(here("data", "gwl", "county", "Sacramento.csv"))

# iterate over files and read them in with a for loop

# list all files we want to read in
files_in <- fs::dir_ls(here("data/gwl/county/"))


# initalize a list of defined length
l <- vector("list", length = length(files_in))
class(l)

# use double bracket notation to access list elements by index 
l[[1]]
l[[2]]

# loop over all files in files_in, and read them into each element of the list 
for(i in seq_along(l)) { # l is object we're going to read them into , could also be 1:3 but this is more consistent
  l[[i]] <- read_csv(files_in[i])
  cat("\nComplete.\n\n")
  
  } 
  
  # we run the loop for i=1, i=2, i=3

#silly loop
for(index in 1:1000){
  cat("We are on index:", index, "of the loop!\n")
}


# A list can hold any R object

my_list <- list(a = 1:3, b = data.frame(x=1, y=2), c="apples")
  

# combine all list elements into a single dataframe 
ldf <- bind_rows(l) # has all of the rows that were in each of the list elements of l 

# split into another list by a column
ldf <- split(ldf, ldf$SITE_CODE)
length(ldf)


# loop over each list element and write a csv file 
fs::dir_create(here("data/gwl/site_code")) # create directory called site_code

# create a vector of filenames to write out
files_out <- glue::glue("{here('data/gwl/site_code', names(ldf))}.csv")


# how we can write each split dataframe of ldf as a separate csv file 
for(i in seq_along(ldf)){
  write_csv(ldf[[i]], files_out[i])
}


```


# lapply
lapply is part of base R and enables functional programming. Simplifies loops with no indexing or initialization

```{r}

files_in

l <- lapply(files_in, read_csv) # same as the for loop above! reads in each file in files_in to list 



```


# write files with mapply
takes more than one input, send files in and out

```{r}

# operates on each set of elements

ldf <- bind_rows(l) #stitches data frames into one

ldf <- split(ldf, ldf$SITE_CODE) # split by site code; there are 723 unique values of site code

length(ldf)

# write out

files_out #output path from above
mapply(function(x, y){write_csv(x,y)} , ldf, files_out) # list objects to send through the factory, this is an anonymous function within mapply (no object in the namespace)

# can also explicitly define the function and use it elsewhere

my_function <- function(x, y) {
  write_csv(x, y)
}

mapply(my_function, ldf, files_out)

# can't do this because write_csv is not set up for working with lists
my_function(ldf, files_out)
class(ldf)
length(files_out)

```



# map

```{r}
# redo what we've already done - read in county csvs 
# like lapply, the first argument to map is the thing you want to operate on/map over 
# tilde ~ is where you start the function you want to map 
# .x is a placeholder for where you put each element you want to map 

l <- map(files_in, ~read_csv(.x))

# as before, let's bind rows, split into dataframes by site code, then write

ldf <- bind_rows(l)

ldf <- group_split(ldf, SITE_CODE) # like split, but doesn't name the list

names(ldf)


# write

walk2(ldf, files_out, ~write_csv(.x, .y)) # like map, but silent; from purrr
# .x is placeholder for first argument, .y for second argument 



```


# map_df 

```{r}

# a common task is to read in dataframes and combine them into one

files_in

map_df(files_in, ~read_csv(.x)) %>%  #read dataframes and combine into one
  group_split(SITE_CODE) %>% # split by SITE_CODE column into separate data frames
  walk2(files_out, ~write_csv(.x, .y)) # walked over the dataframes above and files_out and wrote csvs







```


# map: putting it all together with a modular function 

```{r}

library(tidyverse)
library(here)
library(purrr)

source(here("scripts/functions/f_import_clean_export.R"))


# create a directory to store the results
fs::dir_create(here("results"))


# files in and files out

files_in <- fs::dir_ls(here("data/gwl/county/"))

basename(files_in)

files_out <- here("results", str_replace_all(basename(files_in), ".csv", ".shp")) #change all the filenames from .csv to .shp


walk2(files_in, files_out, ~f_import_clean_export(.x, .y)) # fills results folder with shapefiles for each of the three counties





```


# Advanced spatial and mapmaking

## Install 

```{r}


library(here) # path management
library(tidyverse) # data wrangling & viz
library(purrr) # iteration
library(janitor) # name cleaning
library(glue) # pasting stuff together
library(here) # file paths

# SPATIAL PACKAGES
library(sf) # analysis tools
library(mapview)  # interactive maps!
mapviewOptions(fgb = FALSE) # to save interactive maps

# library(devtools)
# devtools::install_github("UrbanInstitute/urbnmapr")

library(urbnmapr) # county & state boundaries
library(units) # for convert spatial and other units
library(dataRetrieval) # download USGS river data
library(tmap) # mapping
library(tmaptools) # mapping tools
```


## Import spatial data 

```{r}
# get CA
ca <- urbnmapr::get_urbn_map(map="states", sf=TRUE) %>% 
  filter(state_abbv == "CA")
# warning is ok!

# get CA county boundary
ca_co <- urbnmapr::get_urbn_map(map = "counties", sf=TRUE) %>%
  filter(state_abbv == "CA")

mapview(ca)

# shapefile
sac_co <- sf::st_read(here("data/shp/sac/sac_county.shp"))

# check CRS
st_crs(sac_co)$epsg

# match with other data
sac_co <- st_transform(sac_co, st_crs(ca_co))
st_crs(sac_co) == st_crs(ca_co) # should be TRUE!

# make a box around sacramento county
# (a grid with an n=1) for inset
sac_box <- st_make_grid(sac_co, n = 1)

sac_grid_10 <- st_make_grid(sac_co, n=10, square=FALSE) # creates hexagonal grid instead of square

mapview(sac_box) + mapview(sac_grid_10) + mapview(sac_co, color="orange", lwd=2 #line width
                                                  )


```


## Static map with base 'plot()'

```{r}

# if we don't specify the geometry, base plot will plot a map for every attribute in dataset

plot(ca_co$geometry)

# make sure we have all the pieces with a quick test plot
plot(ca$geometry, col = alpha("gray", 0.5), border = "black", lwd=2)



# add layers, use add=TRUE
plot(ca_co$geometry, add = T, border = "purple", col = NA)
plot(sac_co$geometry, add = T, border = "cyan4", col = "skyblue",alpha=0.4, lwd = 2)
plot(sac_box, add = T, border = "orange", col = NA, lwd = 1.4)
plot(sac_grid_10, add=T, border="green", col=NA, lwd=0.5) # add grid over the top



# if something isn't showing up, check projections (iteratively!)
map(list(ca_co, ca, sac_co), ~st_crs(.x)$epsg)


```


## Get groundwater stations (iteratively)

```{r}

# read the stations
gw_files <- list.files(path = here("data/gwl/county"),
                       full.names = TRUE, pattern = "*.csv")

# read all files into dataframes and combine with purrr
gw_df <- map_df(gw_files, ~read.csv(.x))

# the readr package will also do this same thing by default
# when passed a list of files with the same data types
gw_df <- read_csv(gw_files)

# now make "spatial" as sf objects
gw_df <- st_as_sf(gw_df, coords=c("LONGITUDE","LATITUDE"), 
                  crs=4326, remove=FALSE) %>% 
  st_transform(., st_crs(ca_co)) # transform to match ca_co

class(gw_df)

# preview!
mapview(gw_df, zcol="COUNTY_NAME", layer.name="GW Stations") +
  mapview(sac_co, legend=FALSE)



```


## Spatial intersection
Use [ ] to subset or crop our data frame

```{r}

# basic subsetting
gw_df[1, ] # 1st row, all columns

gw_sac_join1 <- gw_df[sac_co, ] # choose rows that match sac_co data frame, which filters down gw_sf to just data points in sac county


plot(sac_co$geometry, col = alpha("forestgreen", 0.6))
plot(gw_sac_join1$geometry, bg = alpha("orange", 0.3), pch=21, add = TRUE)
title("GW Stations (orange) \nthat occur in Sacramento County")


gw_sac_join2 <- st_join(gw_df, sac_co, left=FALSE)
# adds the attributes from the y object to the x object 

mapview(gw_sac_join2, col.region="orange", legend=FALSE) +
  mapview(sac_co, alpha = 0.5, legend = FALSE)



# how to do it using simple filtration and make a new map to look at well use 

gw_sac <- gw_df %>% 
  filter(COUNTY_NAME == "Sacramento")

table(gw_sac$WELL_USE)

mapview(gw_sac, zcol="WELL_USE", layer.name="Well Use")



```


# River data - find nearest USGS station

```{r}

sac_loi <- gw_sac %>% filter(STN_ID %in% c("52418", "5605"))

mapview(sac_co, alpha.regions=0, 
          color="black", lwd=3, legend=FALSE) +
  mapview(sac_loi, col.regions="purple",layer.name="Locations of Interest")

install.packages("dataRetrieval")
library(dataRetrieval)

# Need to convert our locations of interest to WGS84
sac_loi <- sac_loi %>% 
  st_transform(4326)

# now we can go get flowline data!
us_nwis <- map(sac_loi$geometry,
                ~findNLDI(location = .x,
                          nav  = c("UM"), 
                          find = c("nwis", "flowlines"),
                          distance_km = 120)
                )


# we can split these data into separate data frame
# and add them as objects to the .Global environment.
# first add names based on our station IDs:
us_nwis <- set_names(us_nwis, nm=glue("id_{sac_loi$STN_ID}"))

# then split into separate dataframes
# us_nwis %>% list2env(.GlobalEnv)

# Or we can combine with map_df
us_flowlines <- map_df(us_nwis, ~rbind(.x$UM_flowlines))
us_nwissite <- map_df(us_nwis, ~rbind(.x$UM_nwissite))

mapview(sac_loi, col.region="purple", legend = TRUE, 
        cex=3, layer.name="GW LOI") +
  mapview(us_nwissite, col.regions="orange", 
          legend = TRUE, layer.name="UM NWIS Sites") +
  mapview(us_flowlines, color="steelblue", lwd=2, 
          layer.name="UM Flowlines", legend=FALSE)
```
















